Data Management and Domain Adjusting Strategies for Implementing Parallel SPH Method in JPUE Simulation

### 2016-01-06 Paper for submission.tex
fixed all sections

\subfloat[SFCs passing all particles and buckets]{\includegraphics[scale=0.20]{SFC_particles_buckets}}

@@

\subfloat[Domain decomposition based on SFC of buckets]{\includegraphics[scale = 0.22]{SFC_particles_buckets_partition}}

@@

What we need to emphasize here is that in theory motion of particles will destroy locality established based on initial coordinates of particles. However, as particles are moving fairly regularly in a JPUE, the locality of most of particles should still be conserved during simulation.
%If we can design a re-numberring method which is computationally cheap, we can restore the locality.

@@

The most basic data structure of SPH are particle, for problem discription, and bucket, for neighbors searching and domain decomposition. Both are defined as classes in C++. Infomation that contained in particle class can be categorized into six categorise: ID(the key), affiliate(rank of the process that the particle belongs to), primitive variables (variables show up as unknows in governing equations, eg. density, velocity, energy), secondary variables (properties that can be computed from primitive variables, they are stored to avoid repeatedly computing, eg. pressure, temperature ect.), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases ect.) and neighbor infomation (it is a vector of particle keys in our application). Similarly,  Infomation that contained in bucket class can also be categorized into different categorise: ID(the key), affiliate(rank of the process that the bucket belong to), domain information (maximum and mimnum coordinates, boundary infomation), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor infomation (keys of 27 neighbor buckets for three dimension and keys of 9 neighbor buckets for two dimension including its own key) and possessed particles. Objects defined based on these two classes are then accesse through hash tables.

@@

As discussed at the beginning of this section, implementation of SPH in more realistic scenarios requires dynamic memory management and flexible data access. One of the fundemental data structures that satisfy such requirement is hash table. Another option is B-tree. We adopt hash table. An implementation of B-tree under a similar situation for mesh based methods can be found in other papers(eg.\cite{patra2003data}).\\

@@

The hash table has O(1) data accessing, adding and deleting properties when there is no confliction. How many conflictions will happen depends on both distribution of keys in the key space and size of the hash table. As the distribution of keys is determined by particles' initial locations (or centroids of buckets), the hash table size is under our control. We can use very large hash table to minimize hash confliction on the expense of sparse data distribution which will lead to high cache missing and low memory efficiency. Or oppositely, we can use smaller hash table size to obtain high memory efficiency on the expense of having more hash conflictions. Abani\cite{patra2003data} did numerical experiments to examine
the effect of the table size on the dierent data management operations.\\

@@

One way to handle hash confliction is using an additional sorted vector attached to the hash table. When several keys hash to the same slot, a vector will be created. The vector is sorted based on keys so that a binary search can be used to find the correct position for adding, deleting or retrievaling . Another option to handle hash confliction is using an additional link list which is more flexible in memory allocation. The average time complexity of binary search is O(log n) while that for linear search based on link list is O(n). However, accessing efficiency of link list is much lower than array based data structure, especially when the link list becomes longger. Choosing of proper way to handle hash confliction greatly depends on the problem itself. For the test problem in this paper, successively adding of particles at the bottom of the eject vent will lead to hash confliction of many particles, which implies a long link list. But considering the very long conflictions only occur on several slots among millions (see figure \ref{fig:Particle_adding_with_link}), we still choose the link list to handle hash confliction. This decision was made based on numerical experiments.\\

@@

\caption{Ununiform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles at a small portition of the domain, pointer to these new particles will be stored in external link list.}

@@

One natural way to hash time-dependent SFC based key $[k,t]$ is to convert the two elements in the key into one number taking $k$ as the higher digit and $t$ as the lower digit of the large number. However, for JPUE simulation, even though ghost particles for wall boundary condition and pressure boundary condition also need to be added during simulation, places for adding of these two types of ghost particles are previsouly empty area. Only ghost particles for eruption boundary condition will be successsively added at the same place: bottom of the vent. That is to say, particles are distributed ununiformly in the $[0,1]^n \times \textbf{T}$ space as shown in figure \ref{fig:Particle_adding_with_link}. To avoid ununiform, very sparse hash table and conserve locality of SFC, we only plug the first number, $k$, of the key, $[k,t]$, into the hashing function, equation (\ref{eq:hash_function}). 

@@

Particles used in the test problem can be categorized into four types based the particle-type-flag: real particle, wall ghost particle, pressure ghost particle and eruption ghost particle. Ghost particles are for imposing of corresponding boundary conditions (see figure \ref{fig:Involved_initial}). As different types of particles involve different amount of computational work, shown by table \ref{tab:Computational_cost_particles} and table \ref{tab:Computational_cost_steps}, we assign different work load weight for different types of particles based on profilling data. Instead of simply using number of contained particles as work load for bucket, work load of each bucket is determined by summing up work load weight of all particles within the bucket. The SFC sequence passing through centroids of all buckets now becomes a weighted sequence. Domain decomposition will conducted based on the weighted SFC of buckets.\\

@@

Domain decomposition will be conducted based the weighted SFC of buckets. Figure \ref{fig:SFC_domain_decomposition} shows how domain is decomposed based on partition of SFC of buckets. The particles are automatically split into several groups along with buckets that contain them. As SFC of buckets is a curve in the three dimensional space, partition of this curve will automatically lead to 3D domain decomposition. A 2D domain decomposition based on SFC of footprint buckets projected by three dimensional buckets was adopted by Dinesh\cite{kumar2013parallel}. A comparision of the scalability of these two schemes (see figure \ref{fig:2D_vs_3D_efficiency}) confirms that 3D domain decomposition is a better choice when processes number is larger. 

@@

partitions. A set of “guest” particles and buckets are used to synchronize data across partitions. To minimize communications, data is synchronized only where needed, using non-blocking MPI communications. 

@@

As a Lagrangian method, SPH is able to automatically adjust computational domain as the postition of the dicretization points are updated at every time step. However, for JPUE simulation, where some fluid ejects into stationary fluid and get mixed due to turbulence, the domain-adjusting feature of SPH will gone. Because the whole domain, which occupied by stationary fluid before ejected fluid reaching there, has to be discretized at the very beginning of simulation. A lot of CPU time will be spent on computing of "stationary" particles. It is pure wasting of computational resources. If simulating of stationary particls can be avoided, the computational cost will be reduced greatly. To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation. We propose a simple strategy to add such feature in our code with low computational cost. We add a scan function to monitor the most outside layer of the domain. When the ejected fluid reaches the boundary of the current domain, ghost particles (for pressure boundary condition) will be turn to real particles and then add new ghost particles for pressure boundary condition. The original work flow (see figure \ref{fig:Work_flow}) is modified to enable such feature (see figure \ref{fig:Work_flow_adjust}).\\

@@

The start point of adding adjusting feature is adding a flag, which we refer to as involved-flag, into particle class. Particles are categorized into three groups based on the value of the involved-flag: involved (involved-flag=2), potential involved (involved-flag=1), and not involved(involved-flag=0). The involved particles are particles that have already been involved in eject mixing. The potential involved particles are particles that have not been involved in eject mixing but adjecent to involved particles. So they will be involved in the near future. These not involved particles are particles that far away from the eject source and are still at the initial state. For these not involved particles, it is meaningless to update its physical properties. As a consequence, searching for neighbors and data communitation for these particles are also unnecessary. That is to say, only potential involved and involved particles need to be simulated.
As simulation goes, the ejected fluid will reach larger area and more and more particles will be influenced. When originally stationary air is influenced by erupted material, the mass fraction of the erupted material will increase from zero to a positive value. So we can determine whether a particle is involved or not based on whether the mass fraction of that particle is larger than a given threshold (10e-5 in our simulation). Other physical properties, such velocity, can also serve as alternative "swich criteria".\\

@@

\caption{Domain adjusting based on Involved flag of particles}

@@

A has-involved-flag is added to bucket class, too. Buckets are then categorized into 3 types based on particles they contain: has involved (has-involved-flag=3), has potential involved(has-involved-flag=1) and has no involved (has-involved-flag=0). Bucket that has any involved particle will be set to be has involved (has-involved-flag=3). And its neighbor buckets that do not contain any involved particle will be set to be has potential involved. 
All particles in has involved buckets or has potential involved buckets will be set to be potential involved except for involved particles. There are two situations that will switch a has potential involved bucket to be a has involved bucket. First, when a involved particle enters a has potential involved bucket. Second, when mass fraction of a particle (It should be a potental involved particle) exceeds the threshold, this particle will turn to a involved particles and if this is the first involved particle in the bucket, the bucket will turn from a has potential involved bucket to a has involved bucekt.\\
The most outside buckets layer of all has potential involved buckets will be scaned every time after updating. If any of these buckets becomes has involved, the domain will be enlarged by turnning the pressure ghost particles to real particles and switching involved-flag from 0 to 1. Also, the buckets that originally contain pressure ghost particles will become has potential involved. New pressure ghost particles and wall ghost particles will be added around the adjusted domain.

@@

As we are targeting at developing data management and paralell strategies for more complicated implementations of SPH which demmand quick and flexible data access, delete and add. The test problem should have such demand. Volcanic eruption which is essentially a mutilple phases, turbulent, ejection mixing flow accompanied with microphysics processes requires more flexibile data management. We adopt a two phase volcanic plume model\cite{suzuki2005numerical} as our test problem. In this model, one phase is air while another phase is ejected material.

@@

\dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) = 0 \label{eq:gov-cs-rho}
\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot (\rho \xi \textbf{v}) = 0 \label{eq:gov-cs-ks}
\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot (\rho \textbf{v} \textbf{v} + p\textbf{I}) = \rho \textbf{g} \label{eq:gov-cs-v}
\dfrac{\partial \rho E}{\partial t} + \nabla \cdot [(\rho E + p )\textbf{v}] = \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}

@@

Where, $C_v$ is specific heat with constant volume, $n$ is mass fraction, $R$ is gas constant. The subscription $m$ represents mixture of ejected material and air, $s$ is solid portion in ejected material, $g$ is gas portion in the ejected material and $a$ is air.\\

@@

Velocity is zero for non-slip wall boundary. We assume the boundary to be adiabatic and the heat flux is zero on the bounday. The pressure of the surrounding atmosphere is specified on pressure boundaries. Except for the pressure, density, velocity, and energy will depend on the solution. We adpot ghost particles to impose these different type of boundary conditions (see figure \ref{fig:Domain_adjusting}). 

@@

There are several review papers \cite{monaghan1992smoothed, monaghan2005smoothed, price2012smoothed, rosswog2009astrophysical, monaghan2012smoothed} which provide a pretty comprehensive view over SPH. We will not cover these basic theory of SPH in this paper.

@@

Free surface flow are in nature turbulent. We adopt $SPH-\varepsilon$ method developed by Monaghan\cite{monaghan2011turbulence} to capture turbulence in the plume. This will result to a filtered velocity for position updating and additional turbulent terms in momentum and energy eqution.

@@

\caption{Excuting time of test case 1 and test case 2}

@@

\caption{Influnece of total work load on strong scalability}

@@

%Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. 
The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0m, 6000m]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2. The computational work load of test case 2 is 8 times of that of the test case 1. The simulations run for 20s physical time. The consumed time and speed up are shown in figure \ref{fig:2cases_time} and \ref{fig:2cases_efficiency}. Linear speed up is observed when number of processes is smaller than 16. Test case2 shows better speed up than test case1 which implies that the overhead of strong scalability can be increased by increasing total amount of work load. We also compared the performance of 2D domain decomposition and 3D domain decomposition. As shown by figure \ref{fig:2D_vs_3D_efficiency}, 2D domain decomposition shows a little bit better speed up than 3D domain decomposition for 32 processese. This can be explained by the fact that 2D domian decomposition can get more regular subdomain and as a results, need less communication. When number of processes increase, it becomes harder for 2D decomposition to balance workload, that is why 3D domain decomposition has better speed up for processes number equals to 64 and 128. The weak scalability test is conducted with the same initial domain and various smoothing length. Each simulation runs for 400 loops. The average number of real particles of each process keep constant. As shown in figure \ref{fig:weak_loglog}, simulation time increases pretty slowly when number of processes is larger than 132. the relative sharp increase in simulation time from 32 processes to 132 processes might cause by a increase of number of neighbor processes. 
%That can be explained by the fact that the total number of footprint buckets in x-y plane is 64 in the test case.
%When we request 128 processes, 64 processes will be totally idle at the beginning. As simulation goes, the domain will expand and more and more idle processes will start getting tasks. That's why we still get some speed up when number of process is 128. 3D domain decomposition is able to divide the work load in a more flexible way and achieve better strong scalability.
\section{Conclusion}
We developed data management strategies for parallel implementation of SPH method using MPI standard to simulate complicated problems, such as JUPE, which requires flexible and fast data retrievalling, adding and deleting. Neighbors searching and domain decomposition is based on background grid which overlaps the domain and keep stationary during simulation. SFC based index scheme, which provides a global numberring methodology which is purely coordinates dependent, is adopted to give each bucket an unique identifier. A time dependent key which is also based on SFC is used as identifier for particle. 
Hashtables with external link list are adopted for accessing particles and buckets data. Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimize interval. The performance of the code was further improved to several times faster by adjusting computational domain according to progress of simulation. 
Scalability tests on our code shew acceptable strong scalability and good weak scalability. 3D domain decomposition provided better strong scalability than 2D domain decomposition when number of processes is larger.
% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captions off option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

>>>>

\subfloat[SFCs passing all particles and buckets]{\includegraphics[scale=0.15]{SFC_particles_buckets}}

@@

\subfloat[Domain decomposition based on SFC of buckets]{\includegraphics[scale = 0.16]{SFC_particles_buckets_partition}}

@@

We need to emphasize here is that in theory motion of particles will destroy locality established based on initial coordinates of particles. However, as particles are moving fairly regularly in a JPUE, the locality of most of particles is largely conserved during simulation. We will revisit this question for more chaotic flows and 
 design a re-numbering method which is computationally cheap, to restore the locality.

@@

The most basic data structure of SPH are particle, for problem description, and bucket, for neighbors searches and domain decomposition. Both are defined as classes in C++. Information that is contained in particle class can be categorized into six categories: ID(the key), affiliate(rank of the process that the particle belongs to), primitive variables (variables show up as unknowns in governing equations, e.g. density, velocity, energy), secondary variables (properties that can be computed from primitive variables, they are stored to avoid repeated computations eg. pressure, temperature), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases  ) and neighbor information (it is a vector of particle keys in our application). Similarly,  Information that is contained in  a bucket class can also be categorized into different categories: ID(the key), affiliate(rank of the process that the bucket belong to), domain information (maximum and minimum coordinates, boundary information), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor information (keys of 27 neighbor buckets for three dimension and keys of 9 neighbor buckets for two dimension including its own key) and owned particles. Objects defined based on these two classes are then accessed through hash tables.

@@

As discussed at the beginning of this section, implementation of SPH in more realistic scenarios requires dynamic memory management and flexible data access. One of the fundamental data structures that satisfy such requirement is hash table. Another option is B-tree. We adopt hash table. An implementation of B-tree under a similar situation for mesh based methods can be found in other papers (eg.\cite{patra2003data}).\\

@@

The hash table has O(1) data access, adding and deleting properties when there is no conflict. How many conflicts will happen depends on both distribution of keys in the key space and size of the hash table. As the distribution of keys is determined by particles' initial locations (or centroids of buckets), the hash table size is under our control. We can use a very large hash table to minimize hash conflicts at the expense of sparse data distribution which will lead to high cache misses and low memory efficiency. Or alternately, we can use smaller hash table size to obtain high memory efficiency on the expense of having more hash conflicts. Patra et al.\cite{patra2003data} did numerical experiments to examine
the effect of the table size on the different data management operations.\\

@@

One way to handle hash conflicts is by using an additional sorted vector attached to the hash table. When several keys hash to the same slot, a vector will be created. The vector is sorted based on keys so that a binary search can be used to find the correct position for adding, deleting or retrieving. Another option to handle hash conflicts is using an additional link list which is more flexible in memory allocation. The average time complexity of binary search is O(log n) while that for linear search based on link list is O(n). However, accessing efficiency of link list is much lower than array based data structure, especially when the link list becomes longer. Choosing of proper way to handle hash conflicts greatly depends on the problem itself. For the test problem in this paper, successively adding particles at the bottom of the eject vent will lead to hash conflicts of many particles, which implies a long link list. But considering the very long conflicts only occur on several slots among millions (see figure \ref{fig:Particle_adding_with_link}), we still choose the link list to handle hash conflicts. This decision was made based on numerical experiments.\\

@@

\caption{Non-uniform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles at a small portion of the domain, pointer to these new particles will be stored in external link list.}

@@

One natural way to hash time-dependent SFC based key $[k,t]$ is to convert the two elements in the key into one number taking $k$ as the higher digit and $t$ as the lower digit of the large number. However, for JPUE simulation, even though ghost particles for wall boundary condition and pressure boundary condition also need to be added during simulation, places for adding of these two types of ghost particles are previsouly empty area. Only ghost particles for eruption boundary condition will be successively added at the same place: bottom of the vent. That is to say, particles are distributed non-uniformly in the $[0,1]^n \times \textbf{T}$ space as shown in figure \ref{fig:Particle_adding_with_link}. To avoid non-uniform, very sparse hash table and conserve locality of SFC, we only plug the first number, $k$, of the key, $[k,t]$, into the hashing function, equation (\ref{eq:hash_function}). 

@@

Particles used in the test problem can be categorized into four types based the particle-type-flag: real particle, wall ghost particle, pressure ghost particle and eruption ghost particle. Ghost particles are for imposing of corresponding boundary conditions (see figure \ref{fig:Involved_initial}). As different types of particles involve different amount of computational work, shown by table \ref{tab:Computational_cost_particles} and table \ref{tab:Computational_cost_steps}, we assign different work load weight for different types of particles based on profilling data. Instead of simply using number of contained particles as work load for bucket, work load of each bucket is determined by summing up work load weight of all particles within the bucket. The SFC sequence passing through centroids of all buckets now becomes a weighted sequence. Domain decomposition  is conducted based on the weighted SFC of buckets.\\

@@

%Domain decomposition will be conducted based the weighted SFC of buckets.
 Figure \ref{fig:SFC_domain_decomposition} shows how domain is decomposed based on partition of SFC of buckets. The particles are automatically split into several groups along with buckets that contain them. As SFC of buckets is a curve in the three dimensional space, partition of this curve will automatically lead to 3D domain decomposition. A 2D domain decomposition based on SFC of footprint buckets projected by three dimensional buckets was adopted by Kumar et al \cite{kumar2013parallel}. A comparison of the scalability of these two schemes (see figure \ref{fig:2D_vs_3D_efficiency}) confirms that 3D domain decomposition is a better choice when processes number is larger. 

@@

partitions. A set of  particles and buckets are used to synchronize data across partitions. To minimize communications, data is synchronized only where needed, using non-blocking MPI communications. 

@@

As a Lagrangian method, SPH is able to automatically define and adjust its computational domain as the position of the discretization points are updated at every time step.  The 
collective support of the SPH particles defines the computational domain. However, for JPUE simulation (and many similar/related phenomena), where some fluid ejects into stationary fluid and gets mixed, the domain-adjusting feature of SPH cannot be taken advantage of as all of the stationary fluid must be represented. %Because the whole domain, which occupied by stationary fluid before ejected fluid reaching there, has to be discretized at the very beginning of simulation. 
A lot of CPU time will be spent on computing associated with these stationary particles.  If simulating of stationary particles can be avoided, the computational cost will be reduced greatly. %To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation. 
We propose a simple strategy to add such feature in our code with low computational cost. We add a scan function to monitor the most outside layer of the domain. When the ejected fluid reaches the boundary of the current domain, ghost particles (for pressure boundary condition) will be turn to real particles and then add new ghost particles for pressure boundary condition. The original work flow (see figure \ref{fig:Work_flow}) is modified to enable such a feature (see figure \ref{fig:Work_flow_adjust}).\\

@@

This is implemented by using a  involved-flag, in the particle class. Particles are categorized into three groups based on the value of the involved-flag: involved (involved-flag=2), potentially involved (involved-flag=1), and not involved (involved-flag=0). The involved particles are particles that have already been affected by the mixing. The potentially involved particles are particles that have not been involved in mixing but are adjacent to involved particles and will thus be be involved in the near future. %These not involved particles are particles that far away from the eject source and are still at the initial state. For these not involved particles, it is meaningless to update its physical properties. As a consequence, searching for neighbors and data communitation for these particles are also unnecessary.
All communication and computation associated with uninvolved particles can be ignored. That is to say, only potential involved and involved particles need to be simulated.
As simulation progresses, the ejected fluid will reach larger area and more and more particles will be influenced. When originally stationary air is influenced by erupted material, the mass fraction of the erupted material will increase from zero to a positive value. So we can determine whether a particle is involved or not based on whether the mass fraction of that particle is larger than a given threshold ($10^{-5} $ in our simulation). Other physical properties, such velocity, can also serve as alternative ``swich criteria".\\

@@

\caption{ Adjusting the domain based on Involved flag of particles}

@@

A similar scheme can be deployed for buckets too resulting in a categorization  of buckets. The additional complexity here is that potentially involved ghost particles must also be accounted for.
%A has-involved-flag is added to bucket class, too. Buckets are then categorized into 3 types based on particles they contain: has involved (has-involved-flag=3), has potential involved(has-involved-flag=1) and has not involved (has-involved-flag=0). Bucket that has any involved particle will be set to be has involved (has-involved-flag=3). And its neighbor buckets that do not contain any involved particle will be set to be has potential involved. 
%All particles in has involved buckets or has potential involved buckets will be set to be potential involved except for involved particles. There are two situations that will switch a has potential involved bucket to be a has involved bucket. First, when a involved particle enters a has potential involved bucket. Second, when mass fraction of a particle (It should be a potental involved particle) exceeds the threshold, this particle will turn to a involved particles and if this is the first involved particle in the bucket, the bucket will turn from a has potential involved bucket to a has involved bucekt.\\
%The most outside buckets layer of all has potential involved buckets will be scaned every time after updating. If any of these buckets becomes has involved, the domain will be enlarged by turning the pressure ghost particles to real particles and switching involved-flag from 0 to 1. Also, the buckets that originally contain pressure ghost particles will become has potential involved. New pressure ghost particles and wall ghost particles will be added around the adjusted domain.

@@

Our target here is to develop data management and parallelization strategies for more complicated implementations of SPH which demand quick and flexible data access, deletion and addition. Thus, the test problem used should have such requirements. Simulations of a volcanic eruption, which is essentially a muti-phase, turbulent, ejection mixing flow accompanied with microphysics processes will place these demands on the framework. Thus, we adopt a two phase volcanic plume model\cite{suzuki2005numerical} as our test problem. In this model, one phase is air while another phase is ejected material.

@@

\begin{eqnarray}
\dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) &=& 0 \label{eq:gov-cs-rho} \\
%\end{equation}
%\begin{equation}
\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot (\rho \xi \textbf{v}) &=& 0 \label{eq:gov-cs-ks}\\
%\end{equation}
%\begin{equation}
\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot (\rho \textbf{v} \textbf{v} + p\textbf{I}) &=& \rho \textbf{g} \label{eq:gov-cs-v} \\
%\end{equation}
%\begin{equation}
\dfrac{\partial \rho E}{\partial t} + \nabla \cdot [(\rho E + p )\textbf{v}] &=& \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
\end{eqnarray}

@@

Where, $C_v$ is specific heat with constant volume, $n$ is mass fraction, $R$ is gas constant. The subscript
 $m$ represents mixture of ejected material and air, $s$ is solid portion in ejected material, $g$ is gas portion in the ejected material and $a$ is air.\\

@@

Velocity is zero for non-slip wall boundary. We assume the boundary to be adiabatic and the heat flux is zero on the boundary. The pressure of the surrounding atmosphere is specified on pressure boundaries. Except for the pressure, density, velocity, and energy will depend on the solution. We adpot ghost particles to impose these different type of boundary conditions (see figure \ref{fig:Domain_adjusting}). 

@@

There are several review papers \cite{monaghan1992smoothed, monaghan2005smoothed, price2012smoothed, rosswog2009astrophysical, monaghan2012smoothed} which provide a pretty comprehensive descriptions of SPH.  These  may be consulted for additional detail.%We will not cover the basic theory of SPH in this paper.

@@

Free surface flows are inherently turbulent. We adopt $SPH-\varepsilon$ method developed by Monaghan\cite{monaghan2011turbulence} to capture turbulence in the plume. This will result in a filtered velocity for position updates and additional turbulent terms in momentum and energy equation.

@@

\caption{Execution time of test case 1 and test case 2}

@@

\caption{Influence of total work load on strong scalability}

@@

Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. Two types of processors -- 8 core XXXXX running at YYY
clock speed with ZZZ memory per core and 12 core AAAAA running at BBBB
clock speed with CCC memory per core. Each node comprised of two sockets with two of these processors. Memory and level 3 cache are shared on each node.
An important caveat in interpreting these results is that they were obtained in settings where the processors were obtained under normal scheduling (hence readily available) wherein the nodes may be shared with other users. Thus our results may suffer from  contention for memory, CPU-memory bandwidth and network access. As of this writing, we are waiting for exclusive access which will undoubtedly provide better results. On the other hand, these results are representative of what will be available during production use.

The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0m, 6000m]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2, respectively. The computational work load of test case 2 is 8 times of that of the test case 1. The simulations run for 20s physical time.  Wall clock time and speed up are shown in figure \ref{fig:2cases_time} and \ref{fig:2cases_efficiency}. Linear speed up is observed when number of processes is smaller than 16. Test case2 shows better speed up than test case1 which implies that the overhead of strong scalability can be increased by increasing total amount of work load. We also compared the performance of 2D domain decomposition and 3D domain decomposition. As shown by figure \ref{fig:2D_vs_3D_efficiency}, 2D domain decomposition shows a little bit better speed up than 3D domain decomposition for 32 processes. This can be explained by the fact that 2D domain decomposition can get more regular subdomain and as a results, need less communication. When number of processes increase, it becomes harder for 2D decomposition to balance workload, that is why 3D domain decomposition has better speed up for processes number equals to 64 and 128. The weak scalability test is conducted with the same initial domain and various smoothing length. Each simulation runs for 400 loops. The average number of real particles of each process keep constant at $259000$. As shown in figure \ref{fig:weak_loglog}, simulation time increases pretty slowly when number of processes is larger than 132. the relative sharp increase in simulation time from 32 processes to 132 processes might be caused by an increase of number of neighbor processes. 
%That can be explained by the fact that the total number of footprint buckets in x-y plane is 64 in the test case.
%When we request 128 processes, 64 processes will be totally idle at the beginning. As simulation goes, the domain will expand and more and more idle processes will start getting tasks. That's why we still get some speed up when number of process is 128. 3D domain decomposition is able to divide the work load in a more flexible way and achieve better strong scalability.

\section{Conclusion}
We developed data management strategies for parallel implementation of SPH method using MPI standard to simulate complicated problems, such as JUPE, which requires flexible and fast data retrievalling, adding and deleting. Neighbors searching and domain decomposition is based on background grid which overlaps the domain and keep stationary during simulation. SFC based index scheme, which provides a global numberring methodology which is purely coordinates dependent, is adopted to give each bucket an unique identifier. A time dependent key which is also based on SFC is used as identifier for particle. 
Hashtables with external link list are adopted for accessing particles and buckets data. Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimize interval. The performance of the code was further improved to several times faster by adjusting computational domain according to progress of simulation. 
Scalability tests on our code shew acceptable strong scalability and good weak scalability. 3D domain decomposition provided better strong scalability than 2D domain decomposition when number of processes is larger.
% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captions off option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}
